---
title: "Code1_Import"
author: "Devi States"
date: "February 28, 2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Import income.csv data
+ going up one level to access data in data folder

```{r}
library(readr)
income<-read_csv("../Data/income.csv")

```

#import chicago data

```{r}
chicago<-read.csv("../data/chicago.csv",header = TRUE)
summary(chicago)
```

#Load Iris.csv data

```{r}
iris<-read.csv("../data/iris.csv", sep = ",", header = FALSE)
summary(iris)

```
#Load EPA data by going up one level in the directory to the data folder

```{r}
USEPAdata<-read.csv("../data/USEPAdata.csv")
sum(is.na(USEPAdata))

sum(sapply(USEPAdata,function(x)sum(is.na(x)))>0)
```
#Example of loading data from excel file
###skiping two rows, setting column types(skipping a column mostly NA's ), using code chunk option to turn off the warning for income having none numerica values 
```{r warning =FALSE}
library(readxl)
gapminder_xlsx <- read_excel("../Data/gapminder.xlsx", skip = 2, 
                             col_types = c("text", "numeric", "numeric", "skip", "numeric", "text") )
summary(gapminder_xlsx)

```

