---
title: "Code3_visualization"
author: "Devi States"
date: "March 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load mtcars

```{r}
data(mtcars)
```

##loading ggplot2
```{r}
library(ggplot2)
ggplot(mtcars, aes(x=hp, y=mpg)) +
  geom_point(aes(size=cyl))
  
```

# Load MASS package

```{r}
library(MASS)
data("whiteside")
```

## Plot Gas vs. Temp
```{r}

plot(whiteside$Temp, whiteside$Gas, 
     xlab = "Outside temperature", 
     ylab = "Heating gas consumption")
```

# Plot whiteside data
```{r}
plot(whiteside)
```
# Apply the plot() function to Insul
```{r}
plot(whiteside$Insul)
```

# load Cars93 is in your workspace

```{r}
summary(Cars93)
```

# Plot Max.Price vs. Price as red triangles

```{r}
plot(Cars93$Price, Cars93$Max.Price,
  pch = 17,
  col = 'red')
```

# Use matplot() to generate an array of two scatterplots

```{r}
matplot(
  UScereal$calories, 
  UScereal[,c('protein', 'fat')], 
  xlab = "calories",
  ylab = "")
```

# Use matplot() to generate an array of three scatterplots

```{r}
matplot(
  UScereal$calories, 
  UScereal[,c('protein', 'fat', 'fibre')], 
  xlab = "calories",
  ylab = "")
```

# Use matplot() to generate an array of four scatterplots

```{r}
matplot(
  UScereal$calories, 
  UScereal[,c('protein', 'fat', 'fibre', 'carbo')], 
  xlab = "calories",
  ylab = "")
```

# Plot y1 vs. x1 

```{r}
plot(anscombe$x1, anscombe$y1)
```
# Plot y2 vs. x2

```{r}
plot(anscombe$x2, anscombe$y2)
```

# Plot y3 vs. x3
```{r}
plot(anscombe$x3, anscombe$y3)
```

# Plot y4 vs. x4
```{r}
plot(anscombe$x4, anscombe$y4)  
```

# Plot the raw duration data

```{r}
plot(geyser$duration, main = "Raw data")

```

# Plot the normalized histogram of the duration data

```{r}
truehist(geyser$duration, main = "Histogram")
```

# Plot the density of the duration data

```{r}
plot(density(geyser$duration), main = "Density")
```

# Plot scatterplot

```{r}
plot(Boston$rad, Boston$zn)
```

# Plot sunflower plot
```{r}
sunflowerplot(Boston$rad, Boston$zn)
```
# Call symbols() to create the default bubbleplot

```{r}
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price))
```

# Repeat, with the inches argument specified

```{r}
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price),
        inches = 0.1)

```
# Create a table of Cylinders by Origin
# Create the default stacked barplot
```{r}
tbl <- table(Cars93$Cylinders, Cars93$Origin)
barplot(tbl)

```
# Load the insuranceData package

```{r}
library(insuranceData)
```

# Use the data() function to load the dataCar data frame
```{r}
data(dataCar)
head(dataCar)
```

# Load the lattice package
# Use xyplot() to draw the conditional scatterplot

```{r}
library(lattice)
xyplot(calories ~ sugars | shelf, data = UScereal)
```


