---
title: "Code4_Analaysis"
author: "Devi States"
date: "April 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load cars

```{r cars}
data("cars")

```

## Summary satistics
+ load income from code 1
```{r}
mean(cars$speed,na.rm = TRUE)
mean(cars$speed,na.rm = TRUE, trim = 0.1)
median(cars$speed,na.rm = TRUE)
mad(cars$speed,na.rm = TRUE)
sd(cars$speed,na.rm = TRUE)
IQR(cars$speed,na.rm = TRUE)
```

#Load the dplyr package

```{r}
library("dplyr")

```

# Load the hflights package

```{r}
library("hflights")

```

# Call both head() and summary() on hflights

```{r}
head(hflights)
summary(hflights)
```

# Print out the columns Origin up to Cancelled of hflights
```{r}
select(hflights, 14:19)
```
# Print out a table containing just ArrDelay and DepDelay

```{r}
select(hflights, 12, 13)
```
# All flights that traveled 3000 miles or more

```{r}
filter(hflights, Distance >= 3000)
```
# Arrange according to carrier and decreasing departure delays

```{r}
arrange(hflights, UniqueCarrier, desc(DepDelay))

```

#import chicago data

```{r}
chicago<-read.csv("../data/chicago.csv",header = TRUE)

```
# Summary satistics of chicago

```{r}
summary(chicago)
```
#Summary statistics by stratum

```{r}
chicago <- mutate(chicago, tempcat = factor(1 * (tmpd > 80), labels = c("cold", "hot")))
hotcold <- group_by(chicago, tempcat)
hotcold
```

#Load library reader
#Import income data
```{r}
library(readr)

'C:/Users/13148/Desktop/Aplied Data Science/Week6/income.csv'
income<-read_csv("C:/Users/13148/Desktop/Aplied Data Science/Week6/income.csv")
```
#Calculate mean, trimmed mean (10%), and median for Male workers salary (M_weekly). 
#Make sure to set null value parameter to TRUE (see ?mean).
#Select the statement that best describes your results.
```{r}
mean(income$M_weekly,na.rm = TRUE)
mean(income$M_weekly,na.rm = TRUE, trim = 0.1)
median(income$M_weekly,na.rm = TRUE)  
```



