---
title: "FinalReport"
author: "Devi States"
date: "May 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction
Study shows there are many factora influence life expectancy such as socioeconomic status, income, education, access to healthcare, health behavior, poor nutrition and lack of exercise etc. In this study,there are six research questions to be answered by using gapminder data. 

### Importing data and libraries:

```{r message=FALSE}
library(gapminder)
library(plyr)
library(dplyr)
library(tidyr)
library(corrplot)
library(lattice)
library(ggplot2)
library(purrr)
library(tidyverse)
library(cluster)

```
## Importing data as gapminder.csv file:   

```{r}
library(readr)
gapdata<-read_csv("../Data/gapminder.csv")

```
## Discrption of data, data wrangling and data cleaning:
###Variables
  1. country	
  2. continent	
  3. year	
  4. lifeExp	life expectancy at birth
  5. pop	total population
  6. gdpPercap	per-capita 

### Number of observations:
```{r}
nrow(gapminder)

```

### Number of variables

```{r}
ncol(gapminder)

```

### Basic summary of statistics are shown below:

```{r}
summary(gapminder)

```

### There are not missing values

```{r}
sum(is.na(gapminder))

```

### Type of variables 

```{r}
lapply(gapminder, class)

```
###The total number of countries in the data set is shown below::
```{r}
length(levels(gapminder$country))

```

###Skewed to the left (strong right tail)

```{r}
hist(gapminder$gdpPercap,breaks = 30,main = "Figure1:Histogram of GDP Per Capita", xlab = "GDP Per Capita (Dollars)")

```

###The gapminder data frame range from 1952 to 2007

```{r}
range(gapminder$year)

```


##Dta wrangling: 
### The most ages ranges from 70.8 to 76.7 years old. 
```{r}
gapminder$lifeExpcat <- cut(gapminder$lifeExp,breaks = 10)
table(gapminder$lifeExpcat)

```

###Preprocessing: 
### Data transformation to normalize gdp
```{r}
gapminder$gdpPercaplog <- log(gapminder$gdpPercap)
hist(gapminder$gdpPercaplog,breaks = 30,main = "Figure2:Histogram of log-transformed GDP Per Capita", xlab = "log of GDP Per Capita (Dollars)")

```


##Exploring data:
###Research Question 1: How does gdp relate to life expectancy for 2007?

```{r}
plot(lifeExp~gdpPercaplog,data = gapminder, main = "Figure-3: Life Expectancy and GDP for 2007",
xlab = "Log of GDP", ylab = "Years")
lifegdpmodel <- lm(lifeExp~gdpPercaplog,data = gapminder)
abline(lifegdpmodel, col = "red")
legend(x = "bottomright", lty = 1, col = "red", legend = "regression line")

```

####
```{r}
summary(lifegdpmodel)$r.squared

```

###Research Question 2: What is life expectancy of different continent?

```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) +
  geom_boxplot(outlier.colour = "blue") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 0.25) +
  ggtitle("Figure-4:Life expectancy by continent") +
  theme(plot.title = element_text(hjust = 0.5))
 
```



###Research Question 3:. What is the GDP distribution for different continents in 2007?

```{r}
GDPbyCont <- gapminder %>% 
  filter(year==2007) %>% 
  group_by(continent) %>% 
  summarize(avgGDP = mean(gdpPercap))
ggplot(GDPbyCont, 
  aes(x = continent, y = avgGDP, color = continent)) +
  geom_point(size=10) +
  ggtitle("Figure-5:2007 GDP by Continent") +
  xlab("Continent") + ylab("Avg. GDP per Capita") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

```

###Research Question 4: What are the top 5 countries with largest GDP per capita in 2007?
####Table-1
```{r}
gapminder %>% filter(year==2007) %>%
  arrange(-gdpPercap) %>% 
  top_n(5)

```


###Research Question 5: What are the top 5 countries with higher life expetancy in 2007?
####Table-2
```{r}
gapminder %>% filter(year==2007) %>%
  arrange(-lifeExp) %>% 
  top_n(5)

```


###Cluster analysis:
## Research Question 6: What is grouping in data for 2007?
###processing the data

```{r}
kmeansdata <- subset(gapminder,year == 2007, select = c("lifeExp","gdpPercap","pop", "country", "continent"))
summary(kmeansdata)
kmeansdata_scaled <- scale(kmeansdata[,1:3])


```
###Finding the best number of clusters

```{r}
library(factoextra)
fviz_nbclust(kmeansdata_scaled,kmeans,method = "wss")

```

###Building the model and exploring an output

```{r}
set.seed(1)
kmeansObj <- kmeans(kmeansdata_scaled, centers=4)
kmeansObj
kmeansdata$cluster <- kmeansObj$cluster
table(kmeansObj$cluster)
```
```{r}
print.data.frame(kmeansdata[,c("cluster", "country", "continent")])

```

###Visualizing the clusters

```{r}
clusplot(kmeansdata_scaled,kmeansObj$cluster, main = "Figure-6: clusterplot of Kmeansdata(Scaled)", labels = 4, color = 1)

```

###Research Question 7: What is life expectancy relation to all years and gdp?

###Multiple regression of all the data

```{r}
mreg1 <- lm(lifeExp~year+gdpPercaplog,data = gapminder)
summary(mreg1)

```
### Checking models assumptions
```{r}
par(mfrow = c(2,2))
plot(mreg1)

```

##Results and Conclusion:
Figure-1 shows most couries have lower GDP.Figure-2 demonstrats log transformation effectively normalizes the data. 

###Research Question 1: How does gdp relate to life expectancy for 2007?
Life expectancy is a very important measure when we compare the health of different countries.Figure-3 shows there are a strong positive corrlation between life expectancy and level of income(log of gdp) in 2007. 

###Research Question 2: What is life expectancy of different continent?
Figure-4 provides a boxplot of life expectancy of different continent. The figure shows Africa has th lowest life expectancy and Oceania and Europe have the highest life expectancy. 

###Research Question 3:. What is the GDP distribution for different continents in 2007?
Figure-5 shows the average GDP distribution for different conitnents in 2007. The figure demonstrates Africa has the lowest GDP and Occean and Europ have the highest GDP Per Capita in 2007. 

###Research Question 4: What are the top 5 countries with largest GDP per capita in 2007?
Table-1 shows Norway, Kuwait, Singapore, United States, and Ireland are the top 5 countries with largest GDP per capita in 2007.

###Research Question 5: What are the top 5 countries with higher life expetancy in 2007?
Table-2 provides Norway, Singapore,Ireland, United States, and Kuwait are he top 5 countries with higher life expectancy in 2007.

###Research Question 6: What is grouping in data for 2007?
Figure-6 shows 4 different cluster of kmeansdata. In addition, figure-6 shows China(#4) and India (#4) are strongly separated from other grouping. 

###Research Question 7: What is life expectancy relation to all years and gdp?
Regression analysis shows that life expectancy was significantly positively related to both year and gdp. The model fits was strong within R square of 0.72. Residual dignostics showed there are some problems with model assumptions.





