---
title: "Final_Project"
author: "Devi States"
date: "April 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Description:
### Epicycle of Analysis
#### Stating and refining the question
      1. How does income relate to life expectancy?
      2. Which countries have lower life expectancy
      3. Which countries have higher life expectancy

```{r}

```

#	Exploring the data
# data from gapminder.org
# used to perform data transformation and manipulation
# used for data visualization
#load tidyverse

```{r}
library(gapminder)
library(dplyr) 
library(ggplot2)
#library(tidyverse)

```
#Data summary
```{r}
data("gapminder")
gapminder <- as.data.frame(gapminder)
summary(gapminder)
class(gapminder)
```
# scater plot

```{r}
  plot(life~income,data = gapminder)

```

# Box plot of Country by region for life expectancy

```{r fig.width=6, fig.height=9}
ggplot(gapminder,aes(x = Country, y = life)) +
  geom_boxplot(na.rm = TRUE) +
  facet_wrap(~region,ncol = 1,scales = "free_x")
class(gapminder)
```

#	Building formal statistical models
##Model 1

```{r}
Model1 <- lm(life~Year+region,data = gapminder) 
summary(Model1)
```
#Interpreting the results
Model1 demonstrates There is a signinficant positive relationship bwtween year and life expactancy. In addition, life expactancy differ different regions. 

```{r}

```

#Communicating the results
```{r}

```






